---
title: "Phylogeny and Species Trait Effects on Detectability"
author: "Peter Solymos"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Phylogeny and Species Trait Effects on Detectability}
output: knitr:::html_vignette
---

```{r update-stuff,eval=FALSE,results='hide',echo=FALSE}
devtools::install_github("borealbirds/lhreg")
devtools::build_vignettes("~/repos/lhreg")
```

## Introduction

This document is a supporting material for the manuscript entitled
*Phylogeny and species traits predict songbird detectability* by
Peter Solymos, Steven M. Matsuoka, Diana Stralberg, Erin M. Bayne, and Nicole K. S. Barker.

The **lhreg** R package contains the (1) data, (2) analysis code used in the manuscript,
and (3) code required to summarize the results and produce tables and figures.

The R package is hosted on [GitHub](https://github.com/borealbirds/lhreg),
Please submit issues [here](https://github.com/borealbirds/lhreg/issues).

The package is archived on Zenodo under the DOI [10.5281/zenodo.574886](https://zenodo.org/badge/latestdoi/90998177).

The package can be installed as:

```{r install,eval=FALSE}
devtools::install_github("borealbirds/lhreg")
```

The present document can be viewed as:

```{r vignette,eval=FALSE}
vignette(topic = "lhreg", package = "lhreg")
```

## Components of detectability

The manuscript refer to singing rate ($SR$, here we also refer to it as `phi`) 
and detection distance ($DD$, which we refer to as `tau`),
which are linked to detectability as explained in this section.
Solymos et al. (2013, DOI [10.1111/2041-210X.12106](https://dx.doi.org/10.1111/2041-210X.12106))
describes the mathematical details of singing rate estimation based on
removal sampling and effective detection distance estimation via distance sampling.
These quantities define the probability of individuals of the species
are available for sampling given presence (often referred to as availability, $p$),
and the probability that an individual of that species is being detected given
it produces a cue (often referred to as perceptibility, $q$).

$p=1-exp(-t~SR)$, where $t$ is the duration of the counting period.
$q=DD^2~(1-exp(-r^2 / DD^2))~/~r^2$, where $r$ is the rounting radius (i.e. not
counting individuals beyond this distance from the observer).
Probability of detection is the produc to the two components: $pq$.

## Data and data processing

### Trait variables

The `lhreg_data` is a data frame with all response variables
(`logphi` is log singing rate, `logtau` is log detection distance)
and trait values (unmodified and transformed versions).

```{r trait-data,message=FALSE}
library(lhreg)
data(lhreg_data)
str(lhreg_data)
with(lhreg_data, plot(exp(logphi), exp(logtau),
    cex=logmass*0.5, col=Mig2, pch=c(21, 22)[Hab2]))
legend("topright", bty="n", pch=c(21, 21, 22, 22), col=c(1,2,1,2),
    legend=c("Migratory/Closed", "Resident/Closed",
    "Migratory/Open", "Resident/Open"))
```

### Phylogenetic correlations

This code was used to average the 1000 pseudo-posterior trees from Jetz et al.
(2012) with Ericson backbone to represent the phylogenetic relationships among the species:

```{r phylo-corr,eval=FALSE}
library(ape)
mph <- read.nexus("11960.tre") # 1000 trees with Ericson backbone
lhreg_tree <- consensus(mph)
table(sapply(mph, function(z) length(z$tip.label)))
CORR <- TRUE
vv <- list()
vv[[1]] <- vcv(mph[[1]], corr=CORR)
for (i in 2:length(mph)) {
    v <- vcv(mph[[i]], corr=CORR)
    v <- v[rownames(vv[[1]]), colnames(vv[[1]])]
    vv[[i]] <- v
}
vvv <- v
for (i in 1:length(v)) {
    vvv[i] <- mean(sapply(vv, function(z) z[i]))
}
spp <- intersect(rownames(lhreg_data), rownames(vvv))
vvv <- vvv[spp,spp]
cor_matrix <- as.matrix(nearPD(vvv, corr=TRUE)$mat)
```

The object `cor_matrix` is part of the **lhreg** package,
relative phylogenies can be reconstructed from it:

```{r heatmap}
library(ape)
data(cor_matrix)
str(cor_matrix)
heatmap(cor_matrix)
```

```{r phylo-tree,fig.height=12,fig.width=7}
plot(compute.brlen(vcv2phylo(cor_matrix)), cex=0.5)
```

## Analysis

Linear model were used to screen trait variables, so that we compared
that *full* model with the intercept only *null* model, with or without
phylogenetic correlation.

### Variable screening for SR

```{r screening-sr}
y <- lhreg_data$logphi
m4 <- lm(y ~ logmass + Mig2 + Nesthm + xMaxFreqkHz + Hab4, lhreg_data)
m3 <- lm(y ~ logmass + Mig2 + Nesthm + xMaxFreqkHz + Hab3, lhreg_data)
m2 <- lm(y ~ logmass + Mig2 + Nesthm + xMaxFreqkHz + Hab2, lhreg_data)
m4s <- step(m4, trace=0)
m3s <- step(m3, trace=0)
m2s <- step(m2, trace=0)
AIC(m4,m3,m2,m4s,m3s,m2s)
summary(m2s)
mp <- update(m2s, .~.-Nesthm)

amp <- anova(mp)
round(structure(100 * amp[,"Sum Sq"] / sum(amp[,"Sum Sq"]),
    names=rownames(amp)), 1)

mp0 <- lm(y ~ 1)
summary(mp0) # null model for SR
summary(mp) # full model for SR
## evaluating interactions
mpx <- lm(y ~ logmass + Mig2 + Nesthm + xMaxFreqkHz + Hab2 +
    logmass:xMaxFreqkHz + Hab2:xMaxFreqkHz, lhreg_data)
mpx2 <- step(mpx)
summary(mpx2)
```

### Variable screening for DD

```{r screening-dd}
y <- lhreg_data$logtau
m4 <- lm(y ~ logmass + Mig2 + Nesthm + xMaxFreqkHz + Hab4, lhreg_data)
m3 <- lm(y ~ logmass + Mig2 + Nesthm + xMaxFreqkHz + Hab3, lhreg_data)
m2 <- lm(y ~ logmass + Mig2 + Nesthm + xMaxFreqkHz + Hab2, lhreg_data)
m4s <- step(m4, trace=0)
m3s <- step(m3, trace=0)
m2s <- step(m2, trace=0)
AIC(m4,m3,m2,m4s,m3s,m2s)
summary(m2s)
mt <- update(m2s, .~.-Nesthm)

amt <- anova(mt)
round(structure(100 * amt[,"Sum Sq"] / sum(amt[,"Sum Sq"]),
    names=rownames(amt)), 1)

mt0 <- lm(y ~ 1)
summary(mp0) # null model for DD
summary(mp) # full model for DD
## evaluating interactions
mtx <- lm(y ~ logmass + Mig2 + Nesthm + xMaxFreqkHz + Hab2 +
    logmass:xMaxFreqkHz + Hab2:xMaxFreqkHz, lhreg_data)
mtx2 <- step(mtx)
summary(mtx2)
```

### Mixed model selection

The following models were compared, the prefix `t` and `p`
indicates `tau` (detection distance) and `phi` (singing rate):

1. `M0`, null, no phylo, no traits (`~1`, `lam=0`)
2. `Mp0`, phylo only, $\lambda=1$ (`~1`, `lam=1`)
3. `Ml0`, phylo only, $lambda$ estimated (`~1`, `lam=NA`)
4. `Mx`, full trait model, no phylo (`~.`, `lam=0`)
5. `Mpx`, full trait model, $\lambda=1$ (`~.`, `lam=1`)
6. `Mlx`, full trait model, $lambda$ estimated (`~.`, `lam=NA`)

In the manuscript, we refer to the following models:

1. `M0` as null model $M_{00}$,
2. `Ml0` as phylogeny-only model $M_{\lambda0}$,
3. `Mx` as trait-only model $M_{0\beta}$,
4. `Mlx` as combined model $M_{\lambda\beta}$.

```{r models,eval=FALSE}
met <- "DE" # can use "SANN", "Nelder-Mead" is quickest
x <- lhreg_data
vc <- cor_matrix

## model matrix definitions
X0 <- matrix(1, nrow(x), 1) # null (for both)
colnames(X0) <- "Intercept"
Xt <- model.matrix(mt) # DD
Xp <- model.matrix(mp) # SR

## --- model selection


## tau model selection

tM0 <- lhreg(Y=x$logtau, X=X0, SE=x$logtauSE, V=vc, lambda=0,
    hessian=TRUE, method=met)
tMp0 <- lhreg(Y=x$logtau, X=X0, SE=x$logtauSE, V=vc, lambda=1,
    hessian=TRUE, method=met)
tMl0 <- lhreg(Y=x$logtau, X=X0, SE=x$logtauSE, V=vc, lambda=NA,
    hessian=TRUE, method=met)
tMx <- lhreg(Y=x$logtau, X=Xt, SE=x$logtauSE, V=vc, lambda=0,
    hessian=TRUE, method=met)
tMpx <- lhreg(Y=x$logtau, X=Xt, SE=x$logtauSE, V=vc, lambda=1,
    hessian=TRUE, method=met)
tMlx <- lhreg(Y=x$logtau, X=Xt, SE=x$logtauSE, V=vc, lambda=NA,
    hessian=TRUE, method=met)

## phi model selection

pM0 <- lhreg(Y=x$logphi, X=X0, SE=x$logphiSE, V=vc, lambda=0,
    hessian=TRUE, method=met)
pMp0 <- lhreg(Y=x$logphi, X=X0, SE=x$logphiSE, V=vc, lambda=1,
    hessian=TRUE, method=met)
pMl0 <- lhreg(Y=x$logphi, X=X0, SE=x$logphiSE, V=vc, lambda=NA,
    hessian=TRUE, method=met)
pMx <- lhreg(Y=x$logphi, X=Xp, SE=x$logphiSE, V=vc, lambda=0,
    hessian=TRUE, method=met)
pMpx <- lhreg(Y=x$logphi, X=Xp, SE=x$logphiSE, V=vc, lambda=1,
    hessian=TRUE, method=met)
pMlx <- lhreg(Y=x$logphi, X=Xp, SE=x$logphiSE, V=vc, lambda=NA,
    hessian=TRUE, method=met)

aict <- AIC(tM0, tMp0, tMl0, tMx, tMpx, tMlx)
aict$dAIC <- aict$AIC-min(aict$AIC)

aicp <- AIC(pM0, pMp0, pMl0, pMx, pMpx, pMlx)
aicp$dAIC <- aicp$AIC-min(aicp$AIC)

cbind(aict, t(sapply(list(tM0, tMp0, tMl0, tMx, tMpx, tMlx),
    function(z) z$summary[c("sigma","lambda"), 1])))

cbind(aicp, t(sapply(list(pM0, pMp0, pMl0, pMx, pMpx, pMlx),
    function(z) z$summary[c("sigma","lambda"), 1])))
```


### Profile likelihood for lambda

Profile likelihood was calculated to understand how traits affect the
strength of phylogeny through the variable $\lambda$:

```{r profile-lik,eval=FALSE}
library(pbapply)
library(parallel)

lam <- seq(0, 2, by=0.01)

cl <- makeCluster(4)

tmp <- clusterEvalQ(cl, library(lhreg))

object <- tMl0
clusterExport(cl, "object")
pl_tMl0 <- pbsapply(lam, function(z, ...) profile_lambda1(object, z, ...), cl=cl,
    method=met)

object <- tMlx
clusterExport(cl, "object")
pl_tMlx <- pbsapply(lam, function(z, ...) profile_lambda1(object, z, ...), cl=cl,
    method=met)

object <- pMl0
clusterExport(cl, "object")
pl_pMl0 <- pbsapply(lam, function(z, ...) profile_lambda1(object, z, ...), cl=cl,
    method=met)

object <- pMlx
clusterExport(cl, "object")
pl_pMlx <- pbsapply(lam, function(z, ...) profile_lambda1(object, z, ...), cl=cl,
    method=met)

stopCluster(cl)
```

### Parametric bootstrap based confidence intervals

Well, scratch this. We do nonparametrics (LOO) which should give what we need. See below

```{r par-boot,eval=FALSE}
B <- 100 # number of runs
#met="Nelder-Mead"

#cl <- NULL
cl <- makeCluster(4)

tmp <- clusterEvalQ(cl, library(lhreg))

PB_tM0 <-  parametric_bootstrap(tM0,  nsim=B, seed=1234, cl=cl, method=met)
PB_tMl0 <- parametric_bootstrap(tMl0, nsim=B, seed=1234, cl=cl, method=met)
PB_tMx <-  parametric_bootstrap(tMx,  nsim=B, seed=1234, cl=cl, method=met)
PB_tMlx <- parametric_bootstrap(tMlx, nsim=B, seed=1234, cl=cl, method=met)

PB_pM0 <-  parametric_bootstrap(pM0,  nsim=B, seed=1234, cl=cl, method=met)
PB_pMl0 <- parametric_bootstrap(pMl0, nsim=B, seed=1234, cl=cl, method=met)
PB_pMx <-  parametric_bootstrap(pMx,  nsim=B, seed=1234, cl=cl, method=met)
PB_pMlx <- parametric_bootstrap(pMlx, nsim=B, seed=1234, cl=cl, method=met)

stopCluster(cl)

PBcf_tMl0 <- t(sapply(PB_pMl0, function(z) z$summary[,1]))
summary(PBcf_tMl0)
PBcf_tMlx <- t(sapply(PB_pMlx, function(z) z$summary[,1]))
summary(PBcf_tMlx)

```

### LOO analysis

Leave one out cross-validation was used to see how well we could predict
the values based on data from the other species, traits and phylogeny.
Mean squared error (MSE) and variance components were calculated based on LOO
cross validation:

```{r loo,eval=FALSE}
## makes sense to use lm for EDR LOO
tM0$coef
c(coef(mt0), log(summary(mt0)$sigma))
tMx$coef
c(coef(mt), log(summary(mt)$sigma))

pM0$coef
c(coef(mp0), log(summary(mp0)$sigma))
pMx$coef
c(coef(mp), log(summary(mp)$sigma))

n <- nrow(x)
cl <- makeCluster(4)
tmp <- clusterEvalQ(cl, library(lhreg))

pr_tM0  <- t(pbsapply(1:n, loo1, object=tM0, cl=cl))
pr_tMp0 <- t(pbsapply(1:n, loo2, object=tMp0, cl=cl))
pr_tMl0 <- t(pbsapply(1:n, loo2, object=tMl0, cl=cl))
pr_tMx  <- t(pbsapply(1:n, loo1, object=tMx, cl=cl))
pr_tMpx <- t(pbsapply(1:n, loo2, object=tMpx, cl=cl))
pr_tMlx <- t(pbsapply(1:n, loo2, object=tMlx, cl=cl))

pr_pM0  <- t(pbsapply(1:n, loo1, object=pM0, cl=cl))
pr_pMp0 <- t(pbsapply(1:n, loo2, object=pMp0, cl=cl))
pr_pMl0 <- t(pbsapply(1:n, loo2, object=pMl0, cl=cl))
pr_pMx  <- t(pbsapply(1:n, loo1, object=pMx, cl=cl))
pr_pMpx <- t(pbsapply(1:n, loo2, object=pMpx, cl=cl))
pr_pMlx <- t(pbsapply(1:n, loo2, object=pMlx, cl=cl))

stopCluster(cl)

SSEt <- c(
    tM0 = sum((pr_tM0[,1:2] - tM0$Y)^2),
    tMp0 = sum((pr_tMp0[,1:2] - tMp0$Y)^2),
    tMl0 = sum((pr_tMl0[,1:2] - tMl0$Y)^2),
    tMx = sum((pr_tMx[,1:2] - tMx$Y)^2),
    tMpx = sum((pr_tMpx[,1:2] - tMpx$Y)^2),
    tMlx = sum((pr_tMlx[,1:2] - tMlx$Y)^2))
SSEp <- c(
    pM0 = sum((pr_pM0[,1:2] - pM0$Y)^2),
    pMp0 = sum((pr_pMp0[,1:2] - pMp0$Y)^2),
    pMl0 = sum((pr_pMl0[,1:2] - pMl0$Y)^2),
    pMx = sum((pr_pMx[,1:2] - pMx$Y)^2),
    pMpx = sum((pr_pMpx[,1:2] - pMpx$Y)^2),
    pMlx = sum((pr_pMlx[,1:2] - pMlx$Y)^2))
MSEt <- SSEt / n
MSEp <- SSEp / n
```

Use `save.image()` to save the results.

## Results

Load results and set some values:

```{r load-results}
load(system.file("extdata", "lhreg-results-DE.rda", package = "lhreg"))

Crit <- -0.5*qchisq(0.95, 1)
ltmp <- seq(0, 1, by=0.0001)
## red-yl-blue
Col <- c("#2C7BB6", "#6BAACF", "#ABD9E9", "#D4ECD3", "#FFFFBF", "#FED690",
    "#FDAE61", "#EA633E", "#D7191C")
Col1 <- Col[1]
Col2 <- rgb(171/255, 217/255, 233/255, 0.5) # Col[3]
Col3 <- Col[9]
Col4 <- rgb(253/255, 174/255, 97/255, 0.5) # Col[7]
prt <- exp(pr_tMx)
prp <- exp(pr_pMlx)
```

### Table with estimates and MSE

```{r table-1}
ff <- function(z) {
    zz <- z$summary

    pcut <- function(p) {
        factor(c("***", "**", "*", "+", "ns")[as.integer(cut(p,
            c(1, 0.1, 0.05, 0.01, 0.001, 0), include.lowest=TRUE, right=FALSE))],
            levels=c("***", "**", "*", "+", "ns"))
    }
    structure(sapply(1:nrow(zz), function(i)
        paste0(round(zz[i,1], 3), " (SE +/- ", round(zz[i,2], 3), pcut(zz[i,4]), ")")),
        names=rownames(zz))
}
zzz <- lapply(list(pM0, pMp0, pMl0, pMx, pMpx, pMlx), ff)
m <- matrix("", length(zzz[[6]]), 6)
rownames(m) <- names(zzz[[6]])
colnames(m) <- c("M0", "Mp0", "Ml0", "Mx", "Mpx", "Mlx")
for (i in 1:6) {
    j <- match(names(zzz[[i]]), rownames(m))
    m[j,i] <- zzz[[i]]
}
ii <- grep("(SE +/- NANA)", m,fixed=TRUE)
m[ii] <- gsub("(SE +/- NANA)", "(fixed)", m[ii], fixed=TRUE)
m[m==""] <- "n/a"
m1 <- m

zzz <- lapply(list(tM0, tMp0, tMl0, tMx, tMpx, tMlx), ff)
m <- matrix("", length(zzz[[6]]), 6)
rownames(m) <- names(zzz[[6]])
colnames(m) <- c("M0", "Mp0", "Ml0", "Mx", "Mpx", "Mlx")
for (i in 1:6) {
    j <- match(names(zzz[[i]]), rownames(m))
    m[j,i] <- zzz[[i]]
}
ii <- grep("(SE +/- NANA)", m,fixed=TRUE)
m[ii] <- gsub("(SE +/- NANA)", "(fixed)", m[ii], fixed=TRUE)
m[m==""] <- "n/a"
m2 <- m

m1 <- rbind(m1, df=aicp$df, dAIC=round(aicp$dAIC, 3),
    XV_MSE=round(MSEp, 3))
m2 <- rbind(m2, df=aict$df, dAIC=round(aict$dAIC, 3),
    XV_MSE=round(MSEt, 3))

print.default(m1, quote=FALSE) # SR results
print.default(m2, quote=FALSE) # DD results
```

Note: estimates and p-values for $\lambda$ taken from profile likelihood results
due to issues with boundary condition when traits used for
detection distance model.

### Profile likelihood figures

```{r fig-1,width=14,height=6}
#pdf("Fig1.pdf", width=14, height=6)
op <- par(mfrow=c(1,2))

Res1 <- pl_pMl0
Res2 <- pl_pMlx
yv <- exp(Res1-max(Res1))
plot(lam, yv, type="n", lwd=1, ylim=exp(c(2*Crit, 0)),
    xlab=expression(lambda), ylab="Profile Likelihood Ratio")
tmp1 <- splinefun(lam, yv)(ltmp)
i1 <- tmp1 > exp(Crit)
yv <- exp(Res2-max(Res2))
tmp2 <- splinefun(lam, yv)(ltmp)
i2 <- tmp2 > exp(Crit)
polygon(c(ltmp[i1], rev(ltmp[i1])),
    c(tmp1[i1], rep(-1, sum(i1))),
    col=Col2, border=NA)
polygon(c(ltmp[i2], rev(ltmp[i2])),
    c(tmp2[i2], rep(-1, sum(i2))),
    col=Col4, border=NA)
abline(h=exp(Crit), col="grey")
lines(ltmp, tmp1, col=Col1, lwd=1)
lines(ltmp[i1], tmp1[i1], col=Col1, lwd=3)
lines(rep(ltmp[which.max(tmp1)], 2), c(1, -1), col=Col1)
lines(ltmp, tmp2, col=Col3, lwd=1)
lines(ltmp[i2], tmp2[i2], col=Col3, lwd=3)
lines(rep(ltmp[which.max(tmp2)], 2), c(1, -1), col=Col3)
box()
legend(0.1, 1, bty="n", border=c(Col1, Col3), fill=c(Col2, Col4), pt.cex=2, pt.lwd=2,
    legend=c(expression(M[lambda*0]-SR), expression(M[lambda*beta]-SR)))
round(c(Max_Ml0=ltmp[which.max(tmp1)], CI=range(ltmp[i1])), 3)
round(c(Max_Mlx=ltmp[which.max(tmp2)], CI=range(ltmp[i2])), 3)

Res1 <- pl_tMl0
Res2 <- pl_tMlx
yv <- exp(Res1-max(Res1))
plot(lam, yv, type="n", lwd=1, ylim=exp(c(2*Crit, 0)),
    xlab=expression(lambda), ylab="Profile Likelihood Ratio")
tmp1 <- splinefun(lam, yv)(ltmp)
i1 <- tmp1 > exp(Crit)
yv <- exp(Res2-max(Res2))
tmp2 <- splinefun(lam, yv)(ltmp)
i2 <- tmp2 > exp(Crit)
polygon(c(ltmp[i1], rev(ltmp[i1])),
    c(tmp1[i1], rep(-1, sum(i1))),
    col=Col2, border=NA)
polygon(c(ltmp[i2], rev(ltmp[i2])),
    c(tmp2[i2], rep(-1, sum(i2))),
    col=Col4, border=NA)
abline(h=exp(Crit), col="grey")
lines(ltmp, tmp1, col=Col1, lwd=1)
lines(ltmp[i1], tmp1[i1], col=Col1, lwd=3)
lines(rep(ltmp[which.max(tmp1)], 2), c(1, -1), col=Col1)
lines(ltmp, tmp2, col=Col3, lwd=1)
lines(ltmp[i2], tmp2[i2], col=Col3, lwd=3)
lines(rep(ltmp[which.max(tmp2)], 2), c(1, -1), col=Col3)
box()
legend(0.1, 1, bty="n", border=c(Col1, Col3), fill=c(Col2, Col4), pt.cex=2, pt.lwd=2,
    legend=c(expression(M[lambda*0]-DD), expression(M[lambda*beta]-DD)))
round(c(Max_Ml0=ltmp[which.max(tmp1)], CI=range(ltmp[i1])), 3)
round(c(Max_Mlx=ltmp[which.max(tmp2)], CI=range(ltmp[i2])), 3)

par(op)

#dev.off()
```

### LOO cross validation plots

```{r fig-2,width=14,height=7}
#pdf("Fig2.pdf", width=14, height=7)

op <- par(mfrow=c(1,2))

Max <- 0.7
plot(prp, xaxs = "i", yaxs = "i", type="n",
    ylim=c(0, Max), xlim=c(0, Max),
    xlab="Time-removal SR Estimate", ylab="LOO SR Estimate")
abline(0,1,lty=1, col=1)
abline(0,2,lty=2, col="grey")
abline(0,1/2,lty=2, col="grey")
abline(0,1.5,lty=2, col="grey")
abline(0,1/1.5,lty=2, col="grey")
points(prp, xaxs = "i", yaxs = "i",
    col=c(Col1,Col3)[as.integer(x$Mig2)],
    cex=0.2+2*x$MaxFreqkHz/10)
box()
legend("topleft", pch=21, col=c(Col1,Col3,1,1),
    pt.cex=c(1.5,1.5,2,1), bty="n",
    legend=c("Migrant", "Resident", "Song Picth: High", "Song Pitch: Low"))
text(0.9*Max, 0.05*Max, expression(M[lambda*beta]-SR))

Max <- 2.1
plot(prt, xaxs = "i", yaxs = "i", type="n",
    ylim=c(0, Max), xlim=c(0, Max),
    xlab="Distance-sampling DD Estimate", ylab="LOO DD Estimate")
abline(0,1,lty=1, col=1)
abline(0,2,lty=2, col="grey")
abline(0,1/2,lty=2, col="grey")
abline(0,1.5,lty=2, col="grey")
abline(0,1/1.5,lty=2, col="grey")
points(prt, xaxs = "i", yaxs = "i",
    cex=0.2+2*x$logmass/5,
    col=c(Col1,Col3)[as.integer(x$Hab2)])
box()
legend("topleft", pch=21, col=c(Col1,Col3,1,1), pt.cex=c(1.5,1.5,2,1), bty="n",
    legend=c("Habitat: Closed", "Habitat: Open", "Body Mass: Large", "Body Mass: Small"))
text(0.9*Max, 0.05*Max, expression(M[0*beta]-DD))

par(op)

#dev.off()
```

